### 2024-12-04 paper/paper.qmd
finalize data section

---
title: "Donor Preferences in Canadian Political Financing"
subtitle: "Understanding Donor Support for Ruling vs. Opposition Parties Between 2013 - 2024"
author: 
  - Maria Mangru
thanks: "Code and data are available at: [https://github.com/MariaMangru/Ontario_Political_Donors](https://github.com/MariaMangru/Ontario_Political_Donors)."
date: today
date-format: long
abstract: "This study investigates how the ruling political parties in Canada's federal and Ontario provincial governments affect the donation behaviors of individual contributors. By analyzing donation records from 2013 to 2024, the research explores whether a party being in power influences the total amount of financial support it receives. The findings reveal that opposition parties often receive more donations than those in power, highlighting donor preferences to support challengers. This insight enhances our understanding of political financing and can help parties and policymakers develop more effective fundraising strategies."
format: pdf
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Political donations are an important part of democratic processes,
providing essential financial support for political campaigns and
influencing electoral competition. In Canada, both federal and
provincial elections rely on contributions from individuals and
organizations. This can shape the political landscape and reflect public
support for different parties. Understanding the factors that drive
donation patterns is crucial for ensuring transparency and fairness in
the democratic process and for minimizing the risk of undue influence on
policy-making.

A key question in political finance is whether a party's position in
power affects the amount of financial support it receives. Studies in
other countries, such as the United States, have shown that incumbents
often benefit from more donations, particularly from interest groups
seeking to maintain influence (@fouirnaies2014financial). However, there
has been less focus on whether this holds true in Canada, where the
political system and donation regulations differ significantly. This
paper aims to fill that gap by analyzing how a party's power status
influences donor behaviour in federal and Ontario provincial elections.

The primary estimand of this study is the effect of a political party's
power status on the total amount of financial donations it receives from
individual contributors. By comparing the financial support received by
parties in power versus those in opposition, this research aims to
understand whether being in power affects a party's fundraising success.

Using donation data from 2013 to 2024, this study examines total
donations by year and party, accounting for factors such as election
timing and party size (major party or not). The analysis employs linear
regression models to quantify the relationship between power status and
donation amounts, providing insights into donor motivations and the
strategic considerations of political parties.

The results indicate that opposition parties tend to receive higher
total donations than those in power, suggesting that donors often
prioritize supporting challengers, possibly to influence political
change. This information is valuable not only for political parties
developing fundraising strategies but also for policymakers and
researchers seeking to ensure fairness and accountability in political
financing.

The paper is organized as follows: Section \ref{sec-data} discusses the
data sources and methodology, including measurement and data cleaning
processes. Section \ref{sec-model} presents the regression models used
in the analysis, along with justification and validation. Section
\ref{sec-results} outlines the results, and Section \ref{sec-discussion}
discusses the implications, limitations, and suggestions for future
research. Section \ref{sec-appendix} offers further detailed insights
into the data, modeling approach, and methodology.

# Data {#sec-data}

```{r echo=FALSE, message=FALSE, warning=FALSE}
#### Workspace setup ####
library(dplyr)
library(ggplot2)
library(lubridate)
library(scales)
library(readr)   
library(tidyr)

#### Read data ####
analysis_data <- read_csv("../data/02-analysis_data/analysis_data.csv")

# Separate data into Ontario and Federal datasets
ontario_data <- analysis_data %>% filter(region == "Ontario")
federal_data <- analysis_data %>% filter(region == "Federal")
```

## Overview

The dataset analyzed in this study encompasses political donation records from 2006 to 2024, covering both the federal level and the province of Ontario in Canada. These records were sourced from the @ijf_political_donors Political Donors Dataset, which systematically collects and maintains comprehensive data on political donations across Canadian jurisdictions. The IJF updates this dataset daily by monitoring election agency websites at the federal level and within each province and territory, ensuring the data remains current and reliable. covering both the federal level and the province of
Ontario in Canada.

The analysis was conducted using both R (@citeR) and Python (@python),
leveraging a range of libraries for data manipulation and visualization.
In R, key packages included dplyr (@dplyr), lubridate (@lubridate),
tidyverse (@tidyverse), ggplot2 (@ggplot), readr (@readr), tibble
(@tibble), and rstanarm (@rstanarm) for statistical modeling. In Python,
pandas (@pandas), requests (@requests), and matplotlib (@matplotlib)
were utilized for additional data processing and visualization tasks

Political donations in Canada are subject to legal disclosure
requirements, promoting transparency and accountability within the
democratic process. Both federal and provincial election finance laws
mandate that political parties, candidates, and associated entities
report donations exceeding specific thresholds. These disclosures vary
by jurisdiction in terms of frequency and reporting formats. This regulatory framework ensures that significant
contributions are systematically documented and publicly accessible,
providing a reliable foundation for this analysis.

The period starting from 2006 is particularly significant because it marks the first federal election involving the newly formed Conservative Party of Canada. In 2003, the Canadian Alliance and the Progressive Conservative Party of Canada merged to create the modern Conservative Party. The 2006 election was the first in which this new entity participated, making it a pivotal point for analyzing donation patterns related to this party.

## Measurement

The dataset originates from @ijf_political_donors compilation of
political donation records, which includes donations made to registered
political parties, party leadership contestants, riding associations,
and individual candidates. Donations are tracked across different types
of electoral events, including general elections, by-elections, and
leadership races. The IJF collects this data by monitoring official
elections agency websites at both the federal level and within each
province and territory.

The target population for this analysis includes all political donations
made to registered political parties, party leadership contestants,
riding associations (also known as electoral district or constituency
associations), and individual riding candidates during the specified
time frame. This comprehensive coverage allows for an in-depth
examination of donation patterns in various political contexts. The data
collection methodology ensures that all significant financial
contributions are captured, providing a robust foundation for analyzing
the relationship between political power and financial support.

Data aggregation involves summing the total donations received by each
political party annually, resulting in a dataset where each row
represents the total donations for a specific party in a given year. Key
variables include Political Party (categorizing major and minor
parties), Donation Year (ranging from 2013 to 2024), and Amount Donated
(the monetary value of each donation). Additional binary variables such
as Recipient_in_Power (indicating whether the party was in power during
the donation year), Party_Size (distinguishing major from minor parties
based on legislative representation), and Election_Year (indicating
whether the donation was made during an election year) are constructed
to facilitate regression analyses.

The data is organized into a structured format with each row detailing a
specific donation event, including the party receiving the donation, the
year it was made, the amount donated, and the constructed binary
indicators. By aggregating donations at the party-year level and
incorporating relevant variables, the dataset effectively translates
real-world donation phenomena into analyzable data points, enabling the
investigation of how political power status influences financial support
in Canadian elections.

## Data Cleaning

The raw donation records were carefully preprocessed to ensure data
integrity before analysis. First, the data was imported using the
read_csv function from the readr package. Records with missing or
incomplete information in key variables were excluded to maintain
quality. Political party names were standardized to address
inconsistencies, such as consolidating variations like "Liberal Party of
Ontario" and "Ontario Liberal Party" under a single name. Donations were
then aggregated at the party-year level to analyze total annual
contributions received by each party. Binary variables, including
Recipient_in_Power, Party_Size, and Election_Year, were created to
support regression analysis. To normalize the distribution and handle
zero values, a new variable, Log_Total_Donations, was generated by
applying a logarithmic transformation to Total_Donations + 1. Finally,
the cleaned dataset was cross-validated against the original records to
ensure accuracy. Additional details about the data cleaning process can
be found in Appendix \ref{sec-appendix}.

## Outcome Variables

The primary outcome variable is the Total Amount Donated, representing
the total monetary value of donations received by each political party
in a given year. This continuous variable is essential for assessing the
level of financial support and understanding its relationship with
factors such as power status and party size.

The distribution of total donation amounts varies between the federal
and provincial levels. At the federal level, the average total donations
are higher, reflecting broader donor bases and different legal
contribution limits.

\newpage

```{r fig-avg-donation, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-avg-donation
#| tbl-cap: Figure Showing the Average Donation Amounts Over Years

library(ggplot2)
library(dplyr)
library(scales)

# Group and summarize the data
donations_by_year_ontario <- ontario_data %>%
  group_by(donation_year) %>%
  summarize(
    total_donations = n(),
    total_amount = sum(amount, na.rm = TRUE),
    avg_amount = mean(amount, na.rm = TRUE)
  )

donations_by_year_federal <- federal_data %>%
  group_by(donation_year) %>%
  summarize(
    total_donations = n(),
    total_amount = sum(amount, na.rm = TRUE),
    avg_amount = mean(amount, na.rm = TRUE)
  )

# Ensure years are in correct order
donations_by_year_ontario$donation_year <- as.integer(donations_by_year_ontario$donation_year)
donations_by_year_federal$donation_year <- as.integer(donations_by_year_federal$donation_year)

# Combine datasets and add a "Region" column
donations_combined <- bind_rows(
  donations_by_year_ontario %>% mutate(Region = "Ontario"),
  donations_by_year_federal %>% mutate(Region = "Federal")
)

# Plot using facets
ggplot(donations_combined, aes(x = donation_year, y = avg_amount, group = Region, color = Region)) +
  geom_line() +
  geom_point() +
  facet_wrap(~ Region, ncol = 1, scales = "free_y") +
  labs(
    x = "Year",
    y = "Average Donation Amount ($)"
  ) +
  scale_x_continuous(breaks = seq(2006, 2024, 2)) +
  scale_y_continuous(labels = dollar_format()) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )

```

@fig-avg-donation shows the trend of average donation amounts over the years for both Ontario and Federal levels. The x-axis represents the years from 2006 to 2024, while the y-axis shows the average donation amount in Canadian dollars. The data reveals fluctuations in donation amounts over time, with notable during 2018 for Ontario and 2024 for Federal.

## Predictor Variables

Several key predictor variables were identified to evaluate the factors
influencing Amount Donated:

-   Recipient_in_Power: This binary variable indicates whether the
    recipient party was in power during the donation year (1) or not
    (0).

-   Party_Size: Another binary variable which categorizes parties as
    major (1) or minor (0) based on their representation in the
    legislature. Major parties are expected to receive more donations
    due to greater visibility, established support bases, and broader
    outreach capabilities. This variable helps in distinguishing the
    influence of party prominence on financial support.

-   Election_Year: This binary variable signifies whether the donation
    was made during an election year (1) or not (0). Election years
    often see increased fundraising activities as parties intensify
    their efforts to secure electoral victories. Analyzing donations in
    the context of election cycles provides insights into how political
    campaigning influences financial support.

-   Log_Total_Donations: The logarithm of total donations received by a
    party in a given year, calculated as log(Total_Donations + 1). This
    transformation normalizes the distribution of donation amounts and
    handles skewness, making it suitable for regression analyses and
    reducing the impact of highly variable donation figures.

```{r tbl-fed-donation, echo=FALSE, message=FALSE, warning=FALSE}
#| label: tbl-fed-donation
#| tbl-cap: Table Showing Donations Based on Recipient Power Status at Federal Level

# Load required libraries
library(dplyr)
library(knitr)
library(kableExtra)
library(scales)

## Donations by Year - Federal
donations_in_power_federal <- federal_data %>%
  group_by(recipient_in_power) %>%
  summarize(
    total_donations = n(),
    total_amount = sum(amount, na.rm = TRUE),
    avg_amount = mean(amount, na.rm = TRUE)
  ) %>%
  mutate(
    total_amount = dollar(total_amount), 
    avg_amount = dollar(avg_amount)      
  )

# Render Ontario Donations Table
kable(donations_in_power_federal, 
      col.names = c("Recipient In Power", "Total Donations", "Total Amount", "Average Donation Amount"),
      format = "latex", 
      booktabs = TRUE,  
      align = c("l", "c", "c", "c")) %>%
  kable_styling(latex_options = c("striped", "hold_position"))

```

In @tbl-fed-donation, parties in power (indicated by 1) received both lower total donation amounts and lower average donation amounts compared to opposition parties. This pattern suggests that donors may prefer to support challengers over incumbents.

```{r tbl-ontario-donation, echo=FALSE, message=FALSE, warning=FALSE}
#| label: tbl-ontario-donation
#| tbl-cap: Table Showing Donations Based on Recipient Power Status at Ontario Level

## Donations Based on Recipient Power Status - Ontario
donations_in_power_ontario <- ontario_data %>%
  group_by(recipient_in_power) %>%
  summarize(
    total_donations = n(),
    total_amount = sum(amount, na.rm = TRUE),
    avg_amount = mean(amount, na.rm = TRUE)
  ) %>%
  mutate(
    total_amount = dollar(total_amount), # Format as dollars
    avg_amount = dollar(avg_amount)      # Format as dollars
  )

# Render Ontario Donations Table
kable(donations_in_power_ontario, 
      caption = "Donations Based on Recipient Power Status - Ontario", 
      col.names = c("Recipient In Power", "Total Donations", "Total Amount", "Average Donation Amount"),
      format = "latex", # For PDF/LaTeX output
      booktabs = TRUE,  # Add LaTeX booktabs style
      align = c("l", "c", "c", "c")) %>%
  kable_styling(latex_options = c("striped", "hold_position"))
```

@tbl-ontario-donation also shows a similar trend of parties in power receiving lower total and average donations.

```{r fig-ontario-total-donation, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-ontario-total-donation
#| tbl-cap: Figure Showing Total Donations by Political Party at Ontario Level

library(dplyr)
library(ggplot2)
library(treemapify)
library(scales)

# Summarize data for treemap
donations_by_party_ontario <- ontario_data %>%
  group_by(political_party) %>%
  summarize(
    total_amount = sum(amount, na.rm = TRUE)
  ) %>%
  arrange(desc(total_amount))

# Create the treemap
ggplot(donations_by_party_ontario, aes(
  area = total_amount, 
  fill = political_party, 
  label = paste(political_party, "\n$", scales::comma(total_amount))
)) +
  geom_treemap() +
  geom_treemap_text(
    color = "white", 
    place = "center", 
    grow = FALSE,
    reflow = TRUE,
    size = 8
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```

@fig-ontario-total-donation displays a tree map of total donations received by each political party in Ontario. The size of each rectangle corresponds to the total amount donated to that party. Major parties like the Progressive Conservative Party of Ontario and the Liberal Party of Ontario dominate the donation landscape, reflecting their significant fundraising capabilities.

```{r fig-federal-total-donation, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-federal-total-donation
#| tbl-cap: Figure Showing Total Donations by Political Party at Federal Level

# Load required libraries
library(dplyr)
library(ggplot2)
library(treemapify)

# Summarize data for treemap
donations_by_party_federal <- federal_data %>%
  group_by(political_party) %>%
  summarize(
    total_amount = sum(amount, na.rm = TRUE)
  ) %>%
  arrange(desc(total_amount))

# Create the treemap
ggplot(donations_by_party_federal, aes(
  area = total_amount, 
  fill = political_party, 
  label = paste(political_party, "\n$", scales::comma(total_amount))
)) +
  geom_treemap() +
  geom_treemap_text(
    color = "white", 
    place = "center", 
    grow = FALSE,
    reflow = TRUE,
    size = 8
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```

@fig-federal-total-donation presents a treemap for federal political parties. Similar to the provincial level, major parties such as the Conservative Party of Canada and the Liberal Party of Canada receive the largest amounts in donations. This visualization highlights the prominence of major parties in attracting financial support at the federal level.

# Model {#sec-model}

To analyze the impact of a political party’s power status on donation
patterns within Canadian federal and Ontario provincial elections,
multiple linear regression models were developed. Specifically, separate
models were constructed for Ontario and Federal levels to account for
regional differences in political dynamics and donation behaviors.
Additionally, a focused regression model was created for the
Conservative and Liberal parties to explore the interaction between
power status and party size.

## Model set-up

To analyze the impact of a political party's power status on donation
patterns, we employed linear regression models for both Ontario and
federal levels:

```{=tex}
\[
\log(\text{Total Donations}_i) = \beta_0 + \beta_1 \cdot \text{Recipient\_in\_Power}_i + \beta_2 \cdot \text{Party\_Size}_i + \beta_3 \cdot \text{Election\_Year}_i + \epsilon_i
\]

Where:

- \(\log(\text{Total Donations}_i)\): The logarithm of total donations received by party \(i\) in a given year.\\
- \(\text{Recipient\_in\_Power}_i\): A binary variable indicating if party \(i\) was in power.\\
- \(\text{Party\_Size}_i\): A binary variable indicating if party \(i\) is a major party.\\
- \(\text{Election\_Year}_i\): A binary variable indicating if it was an election year.\\
- \(\epsilon_i\): The error term.
```
For the Conservative and Liberal parties, we included an interaction
term:

```{=tex}
\[
\begin{split}
\log(\text{Total Donations}) = \beta_0 + \beta_1 \cdot \text{Recipient in Power} 
+ \beta_2 \cdot \text{Party} \\
+ \beta_3 \cdot (\text{Recipient in Power} \times \text{Party}) 
+ \beta_4 \cdot \text{Election Year} + \epsilon
\end{split}
\]

Where:
- \(\text{Party}_i\): A binary variable indicating 1 for Conservative, 0 for Liberal.\\
```
Each model uses the logarithm of total donations (Log_Total_Donations)
as the dependent variable to stabilize variance and normalize the
distribution of donation amounts. Predictor variables include binary
indicators for whether the party was in power (In_Power), the size of
the party (Party_Size), and whether the donation was made during an
election year (Election_Year). The specialized model for Conservative
and Liberal parties includes an additional binary variable (Party) and
an interaction term (In_Power \* Party) to capture the combined effect
of power status and party type on donation amounts.

### Model justification

Linear regression was chosen for its simplicity and interpretability.
The log transformation of the dependent variable stabilizes variance and
normalizes the distribution, making it suitable for linear modeling. The
models account for key factors influencing donation amounts, allowing us
to isolate the effect of a party being in power.

Alternative models, such as generalized linear models, were considered
but deemed unnecessary due to the adequacy of linear regression for this
analysis.

### Assumptions and Limitations

The models assume a linear relationship between the predictors and the
logarithm of total donations, ensuring that the effects of the
predictors are additive and proportional. Additionally, they assume that
the residuals are normally distributed (normality of errors) and exhibit
constant variance (homoscedasticity). The models also rely on the
absence of multicollinearity, meaning that the predictors are not highly
correlated, and they assume independence of observations, where each
data point is independent of the others. Potential limitations of these
models include the possibility of unobserved variables influencing
donation amounts and the inherent assumption that past trends can
reliably predict future behavior.

### Model Validation \*\*\*\*\*\*

Model validation was conducted to assess the predictive accuracy and
generalizability of the regression models. This involved evaluating key
metrics such as R-squared (R²), Adjusted R-squared, and Root Mean Square
Error (RMSE) to determine how well the models explain the variance in
Log_Total_Donations and their predictive performance.

For the Ontario Model, the regression analysis yielded an R² of 0.645
and an Adjusted R² of 0.635, indicating that approximately 64.5% of the
variance in log-transformed total donations is explained by the
predictors. The RMSE for this model was calculated to assess the average
prediction error, demonstrating a reasonable level of accuracy.

The Federal Model showed an R² of 0.515 and an Adjusted R² of 0.503,
suggesting that 51.5% of the variance in log-transformed total donations
is accounted for by the predictors. The RMSE value indicated acceptable
predictive performance, though slightly lower than the Ontario model,
reflecting potential differences in donation dynamics at the federal
level.

In the Conservative and Liberal Parties Model, the inclusion of the
interaction term between In_Power and Party resulted in an R² of 0.630
and an Adjusted R² of 0.600. This model effectively captures the
combined effect of power status and party type on donation amounts, with
the interaction term providing significant insights into how being in
power influences donations differently for Conservative and Liberal
parties. The RMSE for this model confirmed its robust predictive
capability, comparable to the other models.

Overall, the validation process confirmed that the linear regression
models are robust and effective in predicting donation amounts based on
the selected predictor variables. The models demonstrate good
explanatory power and reasonable predictive accuracy, making them
suitable for analyzing the factors influencing political donations in
Canada. However, ongoing evaluation and potential model refinement are
recommended to accommodate evolving political and financial landscapes,
ensuring the models remain relevant and accurate over time.

# Results {#sec-results}

Table @tbl-modelsummary presents the summary of key model estimates for
the Ontario, Federal, and Conservative/Liberal models. These models
predict the logarithm of total donations received by political parties
based on whether the party was in power, the size of the party, and
whether the donation occurred during an election year. The
Conservative/Liberal model includes an interaction term to explore how
the effect of being in power varies between these two major parties.

```{r echo=FALSE, message=FALSE, warning=FALSE}
#| label: tbl-modelsummary
#| tbl-cap: "Summary of Key Model Estimates for Ontario, Federal, and Conservative/Liberal Models"
#| echo: false
#| message: false

library(modelsummary)

#### Read data ####
analysis_data <- read_csv("../data/02-analysis_data/analysis_data.csv")

#### Model for Ontario ####

# Filter data for Ontario
ontario_data <- analysis_data %>% 
  filter(region == "Ontario")

# Aggregate total donations per party per year in Ontario
total_donations_ontario <- ontario_data %>%
  group_by(political_party, donation_year) %>%
  summarize(
    Total_Donations = sum(amount, na.rm = TRUE),
    In_Power = first(recipient_in_power),
    Party_Size = first(Party_Size),
    Election_Year = first(Election_Year)
  ) %>%
  ungroup()

# Log-transform the dependent variable to stabilize variance
total_donations_ontario <- total_donations_ontario %>%
  mutate(Log_Total_Donations = log(Total_Donations + 1))  # Adding 1 to avoid log(0)

# Run Regression for Ontario with Controls
model_ontario <- lm(Log_Total_Donations ~ In_Power + Party_Size + Election_Year, data = total_donations_ontario)

#### Model for Federal ####

# Filter data for Federal
federal_data <- analysis_data %>% 
  filter(region == "Federal")

# Aggregate total donations per party per year at Federal level
total_donations_federal <- federal_data %>%
  group_by(political_party, donation_year) %>%
  summarize(
    Total_Donations = sum(amount, na.rm = TRUE),
    In_Power = first(recipient_in_power),
    Party_Size = first(Party_Size),
    Election_Year = first(Election_Year)
  ) %>%
  ungroup()

# Log-transform the dependent variable to stabilize variance
total_donations_federal <- total_donations_federal %>%
  mutate(Log_Total_Donations = log(Total_Donations + 1))  # Adding 1 to avoid log(0)

# Run Regression for Federal with Controls
model_federal <- lm(Log_Total_Donations ~ In_Power + Party_Size + Election_Year, data = total_donations_federal)

# Filter for Conservative and Liberal parties
cl_data <- analysis_data %>%
  filter(political_party %in% c(
    "Conservative Party of Canada",
    "Liberal Party of Canada",
    "Progressive Conservative Party of Ontario",
    "Liberal Party of Ontario"
  ))

# Create variables
cl_data <- cl_data %>%
  mutate(
    In_Power = recipient_in_power,  # Already binary (1/0)
    Party = ifelse(
      political_party %in% c("Conservative Party of Canada", "Progressive Conservative Party of Ontario"),
      1,  # Conservative
      0   # Liberal
    )
  )

# Aggregate total donations per party per year
total_donations_cl <- cl_data %>%
  group_by(political_party, donation_year) %>%
  summarize(
    Total_Donations = sum(amount, na.rm = TRUE),
    In_Power = first(In_Power),
    Party = first(Party),
    Election_Year = first(Election_Year)
  ) %>%
  ungroup()

# Log-transform the dependent variable to stabilize variance
total_donations_cl <- total_donations_cl %>%
  mutate(Log_Total_Donations = log(Total_Donations + 1))  # Adding 1 to avoid log(0)

# Run the regression with interaction term
model_cl <- lm(Log_Total_Donations ~ In_Power * Party + Election_Year, data = total_donations_cl)

# Define the list of models
models <- list(
  "Ontario" = model_ontario,
  "Federal" = model_federal,
  "Conservative/Liberal" = model_cl
)

# Define custom coefficient labels
coef_labels <- c(
  "(Intercept)" = "Intercept",
  "In_Power" = "In Power",
  "Party_Size" = "Party Size",
  "Election_Year" = "Election Year",
  "Party" = "Party (1=Conservative, 0=Liberal)",
  "In_Power:Party" = "In Power * Party"
)

# Generate the regression table without significance stars, rounded to 3 decimals
modelsummary(
  models, 
  coef_map = coef_labels, 
  stars = FALSE,
  fmt = 3,
  gof_omit = "AIC|BIC|Log.Lik.|F|RMSE|R2",
  output = "markdown"
)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-estimates
#| fig-cap: "Coefficient Estimates for Ontario, Federal, and Conservative/Liberal Models"
#| echo: false
#| message: false

library(broom)
library(ggplot2)
library(dplyr)
# Tidy models and combine
tidy_all <- bind_rows(
  tidy(model_ontario) %>% mutate(Model = "Ontario"),
  tidy(model_federal) %>% mutate(Model = "Federal"),
  tidy(model_cl) %>% mutate(Model = "Conservative/Liberal")
)

# Select relevant terms and exclude intercepts
tidy_plot <- tidy_all %>%
  filter(term != "(Intercept)") %>%
  mutate(term = case_when(
    term == "In_Power" ~ "In Power",
    term == "Party_Size" ~ "Party Size",
    term == "Election_Year" ~ "Election Year",
    term == "Party" ~ "Party (1=Conservative, 0=Liberal)",
    term == "In_Power:Party" ~ "In Power * Party",
    TRUE ~ term  # Keeps other terms unchanged
  ))

# Plot coefficients with error bars
ggplot(tidy_plot, aes(x = term, y = estimate, color = Model)) +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), 
                width = 0.2, position = position_dodge(width = 0.5)) +
  labs(
    title = "Coefficient Estimates for Ontario, Federal, and Conservative/Liberal Models",
    x = "Predictors",
    y = "Estimated Coefficients"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

The regression analysis reveals that Party Size is a significant
predictor of total donations across both the Ontario and Federal models,
indicating that larger parties receive more financial support.
Specifically, in the Ontario model, Party Size has a coefficient of
2.302 (p \< 0.001), while in the Federal model, it is 3.524 (p \<
0.001). This positive relationship suggests that party prominence plays
a crucial role in attracting donations.

Conversely, the In Power variable is not statistically significant in
the Ontario (p = 0.322) and Federal (p = 0.529) models, implying that
being in power does not directly influence the total donations received
by a party at these levels. However, in the Conservative/Liberal model,
In Power has a significant positive effect (2.513, p \< 0.001), and the
interaction term In Power \* Party is significantly negative (-4.865, p
\< 0.001). This interaction indicates that while being in power
increases donations for Conservative parties, it decreases donations for
Liberal parties, highlighting a differential impact based on party
affiliation.

The Election Year variable does not significantly predict donation
amounts in any of the models, with p-values exceeding conventional
significance thresholds (Ontario p = 0.616, Federal p = 0.975,
Conservative/Liberal p = 0.325). This suggests that donations are not
substantially influenced by whether they occur during an election year.

Overall, the models explain a moderate portion of the variance in total
donations, with R² values ranging from 0.3046 (Ontario) to 0.5825
(Federal), and the Conservative/Liberal model achieving an R² of 0.4481.
These findings underscore the importance of party size in determining
donation amounts, while the role of power status appears to be
contingent on party affiliation.

# Discussion {#sec-discussion}

## The Influence of Party Size on Donation Amounts {#sec-first-point}

At the heart of politics is finance, as revealed by Traag's paper on the complex contagion of campaign donations @traag2016complex. Their study demonstrates that a majority of campaign contributions in the US originate from a few wealthy individuals through large social networks. With the constant battle between government and corporations, the findings in this study underscore the significant influence of financial power in shaping political dynamics.

The landscape in Canada for political financing is growing each year. Results from studies demonstrate that individual contributors may prefer to support challengers for several reasons. An optimistic but naive explanation may be that individuals hope to foster competitive elections, given that donations depend not on political leaning but on who is currently in power. However, a more realistic explanation for this phenomenon is that some individuals want to make policy changes. It is no secret that politicians, particularly those out of power, often solicit funding and political support in exchange for advancing the interests of their stakeholders @mollona2022ethics}. These two dynamics may be working in tandem to create an effect where parties eager to be elected rely more on funding from external sources, which can switch each election cycle.

Moreover, these trends of political donations fluctuating based on the party in power are not unique to Canada. The study "Financing UK Democracy: A Stocktake of 20 Years of Political Donations" highlights a surge in private political donations, more specifically from superdonors, widening the resource gap between large parties @draca2022financing}. Similar to Canada, the UK also sees large donations from a wealthy few influencing their elections each year. The parallel between Canada and the UK underscores a broader democratic issue of financial power being used to influence political outcomes. Though this concept is not novel, this study contributes to the already growing spotlight on the persistent lack of proper reform in political systems.

## Weaknesses and next steps
This study provides valuable insights into donation behaviours across Canada but it its limitations. First, the analysis is applied to only federal and provincial donation records, overlooking regional variations found in other provinces and territories. The geographical limitations of this study hinder its ability to generalize across the entire country. Moreover, the study only spans from 2006 to 2024 which economic or political events could have influence over the donation patterns that the models do not account for. Additionally qualitative factors such as personal beliefs, specific policy preferences and campaign effectiveness are not directly addressed in this study. Data quality issues such as potential missing values or inconsistencies could also have affected the accuracy of the findings here. Lastly, the linear regression models here assume a linear relationship between predictors and donations which can greatly oversimplify the complexities.

Future research should not only aim to address these limitations, but expand the analysis to other provinces and territories across Canada. This would enhance the generalisability of the results and provide a more nuanced understanding of regional differences. Furthermore, utilizing nonlinear modelling techniques or looking at a broader set of predictor variables such as donor demographics or economic indicators could better capture the nuances of donor behaviour and also mitigate the risk of omitted variable bias. Comparative studies with other democratic countries, beyond Canada, could provide more general perspectives on how other political systems and regulatory frameworks influence or are influenced by donation behaviours. Other valuable areas of future exploration include assessing the impact of media coverage on donation trends. Finally, by addressing these areas, research can build on the current study's findings and contribute to a greater understanding of political donations.

\newpage

\appendix

# Appendix {#sec-appendix}

## Appendix A: Data Cleaning Notes

The dataset used in this analysis underwent a rigorous data cleaning
process to ensure accuracy, consistency, and reliability. The raw
donation records were sourced from the Investigative Journalism
Foundation (IJF), encompassing political donations at both the federal
and Ontario provincial levels from 2013 to 2024. The data cleaning
process involved several critical steps, detailed below.

Initially, the dataset was examined for missing or incomplete values in
key variables such as Amount Donated, Political Party, Donation Year,
and Region. Records with missing values in these essential fields were
excluded from the analysis to maintain data integrity and prevent biases
that could arise from incomplete data entries.

Political party names often appeared in various formats due to
inconsistencies in reporting. To address this, party names were
standardized across all records. For instance, variations such as
"Liberal Party of Ontario" and "Ontario Liberal Party" were consolidated
under the standardized name "Ontario Liberal Party." This
standardization ensured accurate attribution of donations to the correct
parties.

Donor information, including names and locations, was also standardized
where possible. Typographical errors were corrected, and formatting was
made consistent (e.g., "First Name Last Name" instead of "Last Name,
First Name"). However, original donor names were retained in cases where
standardization could introduce inaccuracies, particularly when dealing
with individuals with similar names.

The dataset included both monetary (amount_monetary) and non-monetary
(amount_non_monetary) contributions. A new variable, Amount Donated, was
created by summing these two fields for each record. This provided a
comprehensive measure of the total value of donations received by each
party. All monetary values were converted to a consistent currency
format, and any discrepancies in currency symbols or formats were
corrected to ensure accurate aggregation and comparison of donation
amounts.

To facilitate analysis at the party-year level, individual donation
records were aggregated. Donations were summed for each political party
within each year, resulting in a dataset where each entry represented
the total donations received by a party in a specific year. This
aggregation allowed for the examination of trends over time and
assessment of the impact of factors such as power status and election
years on donation amounts.

Several binary variables were introduced to capture key aspects of the
data:

Recipient_in_Power: Indicates whether the recipient party was in power
during the donation year (1 for in power, 0 for not in power).
Party_Size: Categorizes parties as major (1) or minor (0) based on
legislative representation and historical influence. Election_Year:
Indicates whether the donation was made during an election year (1 for
election year, 0 for non-election year). These variables were crucial
for the regression analysis, allowing for the assessment of the
influence of these factors on total donations.

The distribution of total donation amounts was highly skewed, with a
small number of large donations and many smaller ones. To address this
skewness and stabilize variance, a logarithmic transformation was
applied to the Total Donations variable, creating Log_Total_Donations.
This transformation also helped in handling zero values by adding a
small constant (e.g., 1) before taking the logarithm.

To ensure the accuracy of the cleaned dataset, cross-validation was
performed against original records and official financial statements
where available. This step involved spot-checking aggregated totals and
individual records to confirm that the cleaning and aggregation
processes did not introduce errors.

Some challenges were encountered due to inconsistencies in the raw data,
such as typographical errors, inconsistent reporting formats, and
missing donor types. While efforts were made to standardize and correct
these issues, some limitations remain due to the quality of the original
data sources. In cases where donor type information was missing and
legal restrictions indicated that only individuals could donate (e.g.,
Ontario after 2017), the donor type was inferred as "Individual." While
this assumption is reasonable based on the legal context, it may not
capture all nuances in the data.

\newpage

## Appendix B: IJF Data Collection Methodology

The Investigative Journalism Foundation (IJF) employs a comprehensive
approach to collect and compile political donation data across Canada.
Understanding the IJF's data collection methodology is essential for
evaluating the reliability and scope of the data used in this analysis.

**Data Sources and Acquisition**

Every day, the IJF checks election agency websites at the federal level
and in each province and territory for new political donations data.
Historical data is obtained from a variety of government sources,
including provincial archives, legislative libraries, and elections
agencies, covering the period from 1993 to the present day, though this
time range varies by jurisdiction.

Political parties and candidates are legally required to submit records
of donations, which are maintained and made publicly accessible by
election agencies. The frequency and format of these disclosures differ
among jurisdictions; some require quarterly returns, while others
mandate annual reports. The IJF systematically gathers these records to
create a unified database of political donations.

**Scope of Data Collected**

The IJF's dataset encompasses donations to various political entities,
including:

-   Parties: Registered political parties at the federal and provincial
    levels.

-   Candidates: Individual candidates running for office.

-   Riding Associations: Also known as electoral district or
    constituency associations.

-   Leadership and Nomination Contestants: Individuals participating in
    party leadership races or seeking nomination as candidates.

The dataset includes records from all 13 provinces and territories, as
well as city-specific donations from municipalities like Victoria and
Vancouver.

**Donation Laws and Contribution Limits**

Contribution limits and the types of entities permitted to donate vary
across jurisdictions. For instance, at the federal level, only
individuals who are Canadian citizens or permanent residents can donate,
with maximum annual contributions set at specified amounts. In contrast,
some provinces allow donations from corporations, unions, and other
organizations and may have different or no contribution limits.

**Data Formats and Conversion**

The source data is available in multiple formats, including downloadable
spreadsheets, PDFs, and HTML files. For jurisdictions where data is only
available in PDF format, the IJF employs Optical Character Recognition
(OCR) technology to convert the documents into machine-readable CSV
files. OCR tools such as Adobe Export PDF, Cometdocs, and ABBYY
FineReader are used for this purpose.

**Data Cleaning and Standardization**

Extensive manual cleaning is performed to correct errors arising from
OCR conversion and to standardize the data. Key steps in the data
cleaning process include:

Correcting OCR Errors: Manual verification and correction of misread
characters (e.g., misinterpreted currency symbols or letters).
Standardizing Donor and Party Names: Ensuring consistency in naming
conventions for donors and political parties (e.g., standardizing
"Progressive Conservative Party of Ontario" across all records).
Formatting Dates and Amounts: Standardizing date formats to YYYY-MM-DD
and ensuring monetary values are consistently formatted. The IJF also
amalgamates similar categories in variables such as donor types to
improve data legibility (e.g., merging "Business" and "Corporation" into
a single "Corporation" category). In cases where donor type information
is missing and laws specify that only individuals can donate (e.g.,
Ontario after 2017), the IJF adds "Individual" to the donor type column.

**Limitations and Challenges**

Original records submitted by political entities may contain
typographical errors or inaccuracies, which are reflected in the
dataset. Incomplete or Inconsistent Data: Variations in reporting
requirements and practices across jurisdictions can lead to
inconsistencies in the data. OCR Limitations: Despite manual
verification, OCR technology may not perfectly capture handwritten or
poorly scanned documents, potentially introducing errors. Conclusion

### Idealized Methodology

The idealized methodology for analyzing how a political party's power
status affects the total donations it receives in Canada involves
conducting a comprehensive survey targeting individual political donors
across the country. This methodology aims to gather detailed data on
donor behaviours, motivations, and perceptions, providing deeper
insights into the factors influencing political donations at both
federal and Ontario provincial levels.

**Target Population**

The target population consists of individuals who have made financial
contributions to Canadian federal or provincial political parties,
candidates, riding associations, or leadership contestants between 2013
and 2024. This includes donors from all provinces and territories,
ensuring representation across different regions and political contexts.

**Sampling Method**

A stratified random sampling technique will be utilized to ensure
proportional representation from various demographics and regions. The
sample will be stratified based on:

-   Region: All provinces and territories.

-   Donor Type: Donors to parties in power versus opposition parties.

-   Demographics: Age, gender, income level, education level, and
    political affiliation.

In Ontario and other populous provinces, oversampling will be conducted
to capture more granular data and reduce the margin of error. This
approach ensures that key voter segments are adequately represented,
particularly those in regions with significant political activity.

**Sample Size and Response Rate**

To maintain a national margin of error of ±3% at a 95% confidence level,
a total sample size of approximately 1,500 respondents is required. This
sample size allows for meaningful analysis of subgroups within the data.
Given an estimated response rate of 10%, outreach efforts will target
approximately 15,000 potential respondents across Canada.

**Recruitment Methods**

Respondents will be recruited using a combination of methods to maximize
reach and diversity:

-   Online Panel Providers: Collaborate with established online panel
    providers such as Ipsos, Leger, or Angus Reid to access a pool of
    verified respondents who have consented to participate in surveys.

-   Email Invitations: Where legally permissible and ethically
    acceptable, utilize contact information from public donor records to
    send email invitations directly to known political donors.

-   Social Media Advertising: Use targeted advertisements on platforms
    like Facebook, Twitter, and LinkedIn to reach potential respondents
    interested in political activities.

-   Telephone Outreach: Conduct telephone surveys using interactive
    voice response (IVR) systems and live calls to reach donors who may
    not be active online, particularly older individuals or those in
    rural areas.

To incentivize participation, all respondents will be offered a small
monetary compensation (e.g., a \$10 electronic gift card) or entered
into a national sweepstakes with a chance to win one of ten \$500 cash
prizes.

**Data Collection Methods**

Data collection will be conducted via:

-   Online Surveys utilizing platforms like Qualtrics or SurveyMonkey to
    administer the survey electronically. The survey will be designed to
    be mobile-friendly and accessible to individuals with disabilities.

-   Telephone Surveys for respondents preferring or requiring this
    method, trained interviewers will administer the survey, ensuring
    consistency with the online questionnaire.

The survey is designed to take approximately 10 to 15 minutes to
complete, minimizing drop-off rates while collecting comprehensive data.

**Data Validation Measures**

Several measures will be implemented to ensure data quality and
integrity:

-   Completion time checks will be used monitor the time respondents
    take to complete the survey to identify and exclude those who rush
    through it (speeding).

-   Implement reCAPTCHA technology to guard against bots and automated
    responses.

-   nclude attention-check questions to identify inattentive or
    fraudulent responses.

-   After data collection, apply statistical weighting to adjust for any
    demographic discrepancies, ensuring the sample accurately reflects
    the population of political donors in Canada.

**Budget Allocation**

The budget for this study is allocated as follows:

-   Survey Development: \$8,000

    -   This covers costs associated with designing the survey,
        including question formulation, structuring for clarity and
        flow, pilot testing, and ensuring compliance with ethical
        standards. Additionally, funds account for bilingual translation
        services (English and French).

-   Sampling and Recruitment: \$15,000

-   Details: Expenses related to accessing online panel respondents,
    purchasing targeted advertising on social media, and collaborating
    with panel providers. Funds also cover the services of statisticians
    to ensure the sampling aligns with the population distribution.

-   Respondent Incentives: \$20,000

-   Details: Allocated for compensating participants with \$10
    electronic gift cards or funding the national sweepstakes prizes.
    Incentives are crucial for improving response rates and attracting a
    diverse respondent pool.

-   Data Collection: \$25,000

-   Details: Operational costs for administering the survey, including
    online platform fees, telephone survey expenses, interviewer
    training, and call center operations. This ensures data collection
    is efficient and reaches a broad audience.

-   Data Analysis and Reporting: \$12,000

-   Details: Funds dedicated to analyzing the collected data, including
    hiring data analysts proficient in statistical software, conducting
    comprehensive analyses, and preparing detailed reports and
    visualizations to communicate findings effectively.

-   Ethical Compliance and Data Security: \$5,000

-   Details: Covers the costs of obtaining ethical approvals, ensuring
    compliance with privacy legislation (e.g., Personal Information
    Protection and Electronic Documents Act), and implementing robust
    data security measures to protect respondent information.

-   Total Estimated Budget: \$85,000

Timeline

-   Months 1-2: Survey design, ethical approval processes, and pilot
    testing to refine the questionnaire.

-   Months 3-4: Sampling frame development and recruitment of
    respondents through various channels.

-   Months 5-6: Data collection via online and telephone surveys.

-   Month 7: Data cleaning, validation, and preliminary analysis.

-   Months 8-9: Comprehensive data analysis and report preparation.

-   Month 10: Dissemination of findings through academic publications
    and presentations.

### Idealized Survey Design

Survey Structure

[Purpose of survey]
You can access the survey here: https://docs.google.com/forms/d/e/1FAIpQLSdj2RxS9fjZ_n6Q54zjSyoJ7tAJT7ssuTF3V1tVk6a_Rhyp4w/viewform?usp=sharing

**Section 1: Demographics**

**State of Residence:** - Dropdown list of all provinces and
territories.

**Age:** Multiple-choice options

-   18-24

-   25-34

-   35-44

-   45-54

-   55-64

**Gender:**

-   Male

-   Female

-   Non-binary/Other

-   Prefer not to say.

**Income Level:**

-   Less than \$25,000

-   \$25,000-\$49,999

-   \$50,000 - \$74,999

-   \$75,000 - \$99,000

-   \$100,000 - \$149,000

-   \$150,000 or more

**Education Level:** Highest degree obtained

-   Less than high school

-   High school diploma or equivalent

-   Some college, no degree

-   Associate degree

-   Bachelor's degree

-   Graduate or professional degree

**Political Affiliation:**

-   Conservative

-   Liberal

-   NDP

-   Other (please specify)

**Section 2: Donation Behavior**

**Are you currently registered to donate to political parties?**

-   Yes

-   No

**How frequently do you donate to political parties?**

-   Never

-   Rarely (once a year or less)

-   Occasionally (a few times a year)

-   Regularly (monthly or more)

**Average Donation Amount:**

-   Less than \$200

-   \$200 - \$1,000

-   \$1,001 - \$5,000

-   \$5,001 - \$10,000

-   More than \$10,000

**Preferred Donation Channels:**

-   Online platforms

-   Mail-in donations

-   In-person events

**Section 3: Motivations for Donating**

**What motivates you to donate to a political party?** (Select all that
apply)

-   Support for specific policies or issues

-   Desire to influence election outcomes

-   Loyalty to the party or its leaders

-   Social pressure or community influence

-   Tax benefits

-   Other (please specify)

**Which factors most influence your decision to donate?** (Select all
that apply)

-   Candidate's leadership qualities

-   Party's stance on key issues

-   Party's past performance

-   Recommendations from peers or influencers

-   Media coverage

-   Other (please specify)

**Section 4: Impact of Campaigns and Events**

**Have recent political events influenced your donation behavior?**

-   Yes

-   No

**If yes, please specify the events and how they influenced your
donations.** (Open-ended)

**How do you perceive the effectiveness of political campaigns in
encouraging donations?**

-   Very ineffective

-   Ineffective

-   Neutral

-   Effective

-   Very effective

**Section 5: Feedback and Suggestions**

**What could political parties do to encourage more donations from
supporters?** (Open-ended) -

**Any additional comments or suggestions regarding political
donations?** (Open-ended)

# References