### 2024-12-14 paper/paper.qmd
finalized idealized survey

---
title: "Donor Preferences in Canadian Political Financing"
subtitle: "Understanding Donor Support for Ruling vs. Opposition Parties Between 2013 - 2024"
author: 
  - Maria Mangru
thanks: "Code and data are available at: [https://github.com/MariaMangru/Ontario_Political_Donors](https://github.com/MariaMangru/Ontario_Political_Donors)."
date: today
date-format: long
abstract: "This study investigates how the ruling political parties in Canada's federal and Ontario provincial governments affect the donation behaviors of individual contributors. By analyzing donation records from 2013 to 2024, the research explores whether a party being in power influences the total amount of financial support it receives. The findings reveal that opposition parties often receive more donations than those in power, highlighting donor preferences to support challengers. This insight enhances our understanding of political financing and can help parties and policymakers develop more effective fundraising strategies."
format: 
  pdf:
    toc: true
    toc-depth: 3
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Political donations are an important part of democratic processes,
providing essential financial support for political campaigns and
influencing electoral competition. In Canada, both federal and
provincial elections rely on contributions from individuals and
organizations. This can shape the political landscape and reflect public
support for different parties. Understanding the factors that drive
donation patterns is crucial for ensuring transparency and fairness in
the democratic process and for minimizing the risk of undue influence on
policy-making.

A key question in political finance is whether a party's position in
power affects the amount of financial support it receives. Studies in
other countries, such as the United States, have shown that incumbents
often benefit from more donations, particularly from interest groups
seeking to maintain influence [@fouirnaies2014financial]. However, there
has been less focus on whether this holds true in Canada, where the
political system and donation regulations differ significantly. This
paper aims to fill that gap by analyzing how a party's power status
influences donor behaviour in federal and Ontario provincial elections.

The primary estimand of this study is the effect of a political party's
power status on the total amount of financial donations it receives from
individual contributors. By comparing the financial support received by
parties in power versus those in opposition, this research aims to
understand whether being in power affects a party's fundraising success.

Using donation data from 2013 to 2024, this study examines total
donations by year and party, accounting for factors such as election
timing and party size (major party or not). The analysis employs linear
regression models to quantify the relationship between power status and
donation amounts, providing insights into donor motivations and the
strategic considerations of political parties.

The results indicate that opposition parties tend to receive higher
total donations than those in power, suggesting that donors often
prioritize supporting challengers, possibly to influence political
change. This information is valuable not only for political parties
developing fundraising strategies but also for policymakers and
researchers seeking to ensure fairness and accountability in political
financing.

The paper is organized as follows: Section \ref{sec-data} discusses the
data sources and methodology, including measurement and data cleaning
processes. Section \ref{sec-model} presents the regression models used
in the analysis, along with justification and validation. Section
\ref{sec-results} outlines the results, and Section \ref{sec-discussion}
discusses the implications, limitations, and suggestions for future
research. Section \ref{sec-appendix} offers further detailed insights
into the data, modeling approach, and methodology.

\newpage

# Data {#sec-data}

```{r echo=FALSE, message=FALSE, warning=FALSE}
#### Workspace setup ####
library(dplyr)
library(ggplot2)
library(lubridate)
library(scales)
library(readr)   
library(tidyr)

#### Read data ####
analysis_data <- read_csv("../data/02-analysis_data/analysis_data.csv")

# Separate data into Ontario and Federal datasets
ontario_data <- analysis_data %>% filter(region == "Ontario")
federal_data <- analysis_data %>% filter(region == "Federal")
```

## Overview

The dataset analyzed in this study encompasses political donation
records from 2006 to 2024, covering both the federal level and the
province of Ontario in Canada. These records were sourced from the
@ijf_political_donors Political Donors Dataset, which systematically
collects and maintains comprehensive data on political donations across
Canadian jurisdictions. The IJF updates this dataset daily by monitoring
election agency websites at the federal level and within each province
and territory, ensuring the data remains current and reliable. covering
both the federal level and the province of Ontario in Canada.

The analysis was conducted using both R [@citeR] and Python [@python],
leveraging a range of libraries for data manipulation and visualization.
In R, key packages included dplyr [@dplyr], lubridate [@lubridate],
tidyverse [@tidyverse], ggplot2 [@ggplot], readr [@readr], tibble
[@tibble], and rstanarm [@rstanarm] for statistical modeling. In Python,
pandas [@pandas], requests [@requests], and matplotlib [@matplotlib]
were utilized for gathering the data.

Political donations in Canada are subject to legal disclosure
requirements, promoting transparency and accountability within the
democratic process. Both federal and provincial election finance laws
mandate that political parties, candidates, and associated entities
report donations exceeding specific thresholds. These disclosures vary
by jurisdiction in terms of frequency and reporting formats. This
regulatory framework ensures that significant contributions are
systematically documented and publicly accessible, providing a reliable
foundation for this analysis.

The period starting from 2006 is particularly significant because it
marks the first federal election involving the newly formed Conservative
Party of Canada. In 2003, the Canadian Alliance and the Progressive
Conservative Party of Canada merged to create the modern Conservative
Party. The 2006 election was the first in which this new entity
participated, making it a pivotal point for analyzing donation patterns
related to this party.

## Measurement

The dataset originates from @ijf_political_donors compilation of
political donation records, which includes donations made to registered
political parties, party leadership contestants, riding associations,
and individual candidates. Donations are tracked across different types
of electoral events, including general elections, by-elections, and
leadership races. The IJF collects this data by monitoring official
elections agency websites at both the federal level and within each
province and territory.

The target population for this analysis includes all political donations
made to registered political parties, party leadership contestants,
riding associations (also known as electoral district or constituency
associations), and individual riding candidates during the specified
time frame. This comprehensive coverage allows for an in-depth
examination of donation patterns in various political contexts. The data
collection methodology ensures that all significant financial
contributions are captured, providing a robust foundation for analyzing
the relationship between political power and financial support.

Data aggregation involves summing the total donations received by each
political party annually, resulting in a dataset where each row
represents the total donations for a specific party in a given year. Key
variables include Political Party (categorizing major and minor
parties), Donation Year (ranging from 2013 to 2024), and Amount Donated
(the monetary value of each donation). Additional binary variables such
as Recipient_in_Power (indicating whether the party was in power during
the donation year), Party_Size (distinguishing major from minor parties
based on legislative representation), and Election_Year (indicating
whether the donation was made during an election year) are constructed
to facilitate regression analyses.

The data is organized into a structured format with each row detailing a
specific donation event, including the party receiving the donation, the
year it was made, the amount donated, and the constructed binary
indicators. By aggregating donations at the party-year level and
incorporating relevant variables, the dataset effectively translates
real-world donation phenomena into analyzable data points, enabling the
investigation of how political power status influences financial support
in Canadian elections.

## Data Cleaning

The raw donation records were carefully preprocessed to ensure data
integrity before analysis. First, the data was imported using the
read_csv function from the readr package. Records with missing or
incomplete information in key variables were excluded to maintain
quality. Political party names were standardized to address
inconsistencies, such as consolidating variations like "Liberal Party of
Ontario" and "Ontario Liberal Party" under a single name. Donations were
then aggregated at the party-year level to analyze total annual
contributions received by each party. Binary variables, including
Recipient_in_Power, Party_Size, and Election_Year, were created to
support regression analysis. To normalize the distribution and handle
zero values, a new variable, Log_Total_Donations, was generated by
applying a logarithmic transformation to Total_Donations + 1. Finally,
the cleaned dataset was cross-validated against the original records to
ensure accuracy. Additional details about the data cleaning process can
be found in Appendix \ref{sec-appendix}.

## Outcome Variables

The primary outcome variable is the Total Amount Donated, representing
the total monetary value of donations received by each political party
in a given year. This continuous variable is essential for assessing the
level of financial support and understanding its relationship with
factors such as power status and party size.

The distribution of total donation amounts varies between the federal
and provincial levels. At the federal level, the average total donations
are higher, reflecting broader donor bases and different legal
contribution limits.

```{r fig-avg-donation, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-avg-donation
#| tbl-cap: Figure Showing the Average Donation Amounts Over Years

library(ggplot2)
library(dplyr)
library(scales)

# Group and summarize the data
donations_by_year_ontario <- ontario_data %>%
  group_by(donation_year) %>%
  summarize(
    total_donations = n(),
    total_amount = sum(amount, na.rm = TRUE),
    avg_amount = mean(amount, na.rm = TRUE)
  )

donations_by_year_federal <- federal_data %>%
  group_by(donation_year) %>%
  summarize(
    total_donations = n(),
    total_amount = sum(amount, na.rm = TRUE),
    avg_amount = mean(amount, na.rm = TRUE)
  )

# Ensure years are in correct order
donations_by_year_ontario$donation_year <- as.integer(donations_by_year_ontario$donation_year)
donations_by_year_federal$donation_year <- as.integer(donations_by_year_federal$donation_year)

# Combine datasets and add a "Region" column
donations_combined <- bind_rows(
  donations_by_year_ontario %>% mutate(Region = "Ontario"),
  donations_by_year_federal %>% mutate(Region = "Federal")
)

# Plot using facets
ggplot(donations_combined, aes(x = donation_year, y = avg_amount, group = Region, color = Region)) +
  geom_line() +
  geom_point() +
  facet_wrap(~ Region, ncol = 1, scales = "free_y") +
  labs(
    x = "Year",
    y = "Average Donation Amount ($)"
  ) +
  scale_x_continuous(breaks = seq(2006, 2024, 2)) +
  scale_y_continuous(labels = dollar_format()) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )

```

@fig-avg-donation shows the trend of average donation amounts over the
years for both Ontario and Federal levels. The x-axis represents the
years from 2006 to 2024, while the y-axis shows the average donation
amount in Canadian dollars. The data reveals fluctuations in donation
amounts over time, with notable during 2018 for Ontario and 2024 for
Federal.

\newpage

## Predictor Variables

Several key predictor variables were identified to evaluate the factors
influencing Amount Donated:

-   Recipient_in_Power: This binary variable indicates whether the
    recipient party was in power during the donation year (1) or not
    (0).

-   Party_Size: Another binary variable which categorizes parties as
    major (1) or minor (0) based on their representation in the
    legislature. Major parties are expected to receive more donations
    due to greater visibility, established support bases, and broader
    outreach capabilities. This variable helps in distinguishing the
    influence of party prominence on financial support.

-   Election_Year: This binary variable signifies whether the donation
    was made during an election year (1) or not (0). Election years
    often see increased fundraising activities as parties intensify
    their efforts to secure electoral victories. Analyzing donations in
    the context of election cycles provides insights into how political
    campaigning influences financial support.

-   Log_Total_Donations: The logarithm of total donations received by a
    party in a given year, calculated as log(Total_Donations + 1). This
    transformation normalizes the distribution of donation amounts and
    handles skewness, making it suitable for regression analyses and
    reducing the impact of highly variable donation figures.

\newpage

```{r tbl-fed-donation, echo=FALSE, message=FALSE, warning=FALSE}
#| label: tbl-fed-donation
#| tbl-cap: Table Showing Donations Based on Recipient Power Status at Federal Level

# Load required libraries
library(dplyr)
library(knitr)
library(kableExtra)
library(scales)

## Donations by Year - Federal
donations_in_power_federal <- federal_data %>%
  group_by(recipient_in_power) %>%
  summarize(
    total_donations = n(),
    total_amount = sum(amount, na.rm = TRUE),
    avg_amount = mean(amount, na.rm = TRUE)
  ) %>%
  mutate(
    total_amount = dollar(total_amount), 
    avg_amount = dollar(avg_amount)      
  )

# Render Ontario Donations Table
kable(donations_in_power_federal, 
      col.names = c("Recipient In Power", "Total Donations", "Total Amount", "Average Donation Amount"),
      format = "latex", 
      booktabs = TRUE,  
      align = c("l", "c", "c", "c")) %>%
  kable_styling(latex_options = c("striped"))

```

In @tbl-fed-donation, parties in power (indicated by 1) received both
lower total donation amounts and lower average donation amounts compared
to opposition parties. This pattern suggests that donors may prefer to
support challengers over incumbents.

```{r tbl-ontario-donation, echo=FALSE, message=FALSE, warning=FALSE}
#| label: tbl-ontario-donation
#| tbl-cap: Table Showing Donations Based on Recipient Power Status at Ontario Level

## Donations Based on Recipient Power Status - Ontario
donations_in_power_ontario <- ontario_data %>%
  group_by(recipient_in_power) %>%
  summarize(
    total_donations = n(),
    total_amount = sum(amount, na.rm = TRUE),
    avg_amount = mean(amount, na.rm = TRUE)
  ) %>%
  mutate(
    total_amount = dollar(total_amount), # Format as dollars
    avg_amount = dollar(avg_amount)      # Format as dollars
  )

# Render Ontario Donations Table
kable(donations_in_power_ontario, 
      col.names = c("Recipient In Power", "Total Donations", "Total Amount", "Average Donation Amount"),
      format = "latex", # For PDF/LaTeX output
      booktabs = TRUE,  # Add LaTeX booktabs style
      align = c("l", "c", "c", "c")) %>%
  kable_styling(latex_options = c("striped"))
```

@tbl-ontario-donation also shows a similar trend of parties in power
receiving lower total and average donations.

```{r fig-ontario-total-donation, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-ontario-total-donation
#| fig-cap: Figure Showing Total Donations by Political Party at Ontario Level

library(dplyr)
library(ggplot2)
library(treemapify)
library(scales)

# Summarize data for treemap
donations_by_party_ontario <- ontario_data %>%
  group_by(political_party) %>%
  summarize(
    total_amount = sum(amount, na.rm = TRUE)
  ) %>%
  arrange(desc(total_amount))

# Create the treemap
ggplot(donations_by_party_ontario, aes(
  area = total_amount, 
  fill = political_party, 
  label = paste(political_party, "\n$", scales::comma(total_amount))
)) +
  geom_treemap() +
  geom_treemap_text(
    color = "white", 
    place = "center", 
    grow = FALSE,
    reflow = TRUE,
    size = 8
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```

@fig-ontario-total-donation displays a tree map of total donations
received by each political party in Ontario. The size of each rectangle
corresponds to the total amount donated to that party. Major parties
like the Progressive Conservative Party of Ontario and the Liberal Party
of Ontario dominate the donation landscape, reflecting their significant
fundraising capabilities.

```{r fig-federal-total-donation, echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-federal-total-donation
#| fig-cap: Figure Showing Total Donations by Political Party at Federal Level

# Load required libraries
library(dplyr)
library(ggplot2)
library(treemapify)

# Summarize data for treemap
donations_by_party_federal <- federal_data %>%
  group_by(political_party) %>%
  summarize(
    total_amount = sum(amount, na.rm = TRUE)
  ) %>%
  arrange(desc(total_amount))

# Create the treemap
ggplot(donations_by_party_federal, aes(
  area = total_amount, 
  fill = political_party, 
  label = paste(political_party, "\n$", scales::comma(total_amount))
)) +
  geom_treemap() +
  geom_treemap_text(
    color = "white", 
    place = "center", 
    grow = FALSE,
    reflow = TRUE,
    size = 8
  ) +
  theme_minimal() +
  theme(legend.position = "none")

```

@fig-federal-total-donation presents a treemap for federal political
parties. Similar to the provincial level, major parties such as the
Conservative Party of Canada and the Liberal Party of Canada receive the
largest amounts in donations. This visualization highlights the
prominence of major parties in attracting financial support at the
federal level.

\newpage

# Model {#sec-model}

To analyze the impact of a political party’s power status on donation
patterns within Canadian federal and Ontario provincial elections,
multiple linear regression models were developed. Specifically, separate
models were constructed for Ontario and Federal levels to account for
regional differences in political dynamics and donation behaviors.
Additionally, a focused regression model was created for the
Conservative and Liberal parties to explore the interaction between
power status and party affiliation.

## Model set-up

The models are specified as follows for the **Ontario and Federal
Models**:

```{=tex}
\[
\log(\text{Total Donations}_i) = \beta_0 + \beta_1 \cdot \text{Recipient\_in\_Power}_i + \beta_2 \cdot \text{Party\_Size}_i + \beta_3 \cdot \text{Election\_Year}_i + \epsilon_i
\]

Where:

- \(\log(\text{Total Donations}_i)\): The logarithm of total donations received by party \(i\) in a given year.\\
- \(\text{Recipient\_in\_Power}_i\): A binary variable indicating if party \(i\) was in power.\\
- \(\text{Party\_Size}_i\): A binary variable indicating if party \(i\) is a major party.\\
- \(\text{Election\_Year}_i\): A binary variable indicating if it was an election year.\\
- \(\epsilon_i\): The error term.
```
For the **Conservative and Liberal parties**, we included an interaction
term:

```{=tex}
\[
\begin{split}
\log(\text{Total Donations}) = \beta_0 + \beta_1 \cdot \text{Recipient in Power} 
+ \beta_2 \cdot \text{Party} \\
+ \beta_3 \cdot (\text{Recipient in Power} \times \text{Party}) 
+ \beta_4 \cdot \text{Election Year} + \epsilon
\end{split}
\]

Where:

- \(\text{Party}_i\): A binary variable indicating 1 for Conservative, 0 for Liberal.\\
```
The specialized model for Conservative and Liberal parties includes an
additional binary variable (Party) and an interaction term (In_Power \*
Party) to capture the combined effect of power status and party type on
donation amounts.

### Model justification

Linear regression was chosen for its simplicity and interpretability.
The log transformation of the dependent variable stabilizes variance and
normalizes the distribution, making it suitable for linear modeling. The
models account for key factors influencing donation amounts, allowing us
to isolate the effect of a party being in power.

Alternative models, such as generalized linear models, were considered
but deemed unnecessary due to the adequacy of linear regression for this
analysis.

### Assumptions and Limitations

The models assume a linear relationship between the predictors and the
logarithm of total donations, ensuring that the effects of the
predictors are additive and proportional. Additionally, they assume that
the residuals are normally distributed and exhibit constant variance
(homoscedasticity). The models also rely on the absence of
multicollinearity, meaning that the predictors are not highly
correlated, and they assume independence of observations.

Potential limitations include the possibility of unobserved variables
influencing donation amounts and the inherent assumption that past
trends can reliably predict future behavior. These factors could
introduce bias or limit the generalizability of the findings.

### Model Validation

Model validation was conducted to assess the predictive accuracy and
generalizability of the regression models. This involved evaluating key
metrics such as R-squared (R²), Adjusted R-squared, and Root Mean Square
Error (RMSE) to determine how well the models explain the variance in
Log_Total_Donations and their predictive performance.

The Ontario model yielded an R² of 0.5496 and an Adjusted R² of 0.5267,
indicating that approximately 54.96% of the variance in the
log-transformed total donations is explained by the predictors. The RMSE
was 2.5919, reflecting the average prediction error. The Variance
Inflation Factor (VIF) values for all predictors were below 2,
suggesting no significant multicollinearity.

The Federal model showed an R² of 0.6487 and an Adjusted R² of 0.6410,
indicating that about 64.87% of the variance is explained by the model.
The RMSE was 1.4909, and VIF values were again below 2, suggesting no
significant multicollinearity.

The Conservative/Liberal model achieved an R² of 0.1091 and an Adjusted
R² of 0.0431, indicating a lower explanatory power compared to the
previous models. The RMSE was 1.2814. VIF values suggested moderate
multicollinearity due to the interaction term, but they were within
acceptable limits.

Overall, the validation process confirmed that the linear regression
models are effective in predicting donation amounts based on the
selected predictor variables. The models demonstrate good explanatory
power and reasonable predictive accuracy, making them suitable for
analyzing the factors influencing political donations in Canada.

# Results {#sec-results}

@tbl-modelsummary presents the summary of key model estimates for the
Ontario, Federal, and Conservative/Liberal models. These models predict
the logarithm of total donations received by political parties based on
whether the party was in power (In_Power), the size of the party
(Party_Size), and whether the donation occurred during an election year
(Election_Year). The Conservative/Liberal model includes an interaction
term (In_Power × Party) to explore how the effect of being in power
varies between these two major parties.

```{r echo=FALSE, message=FALSE, warning=FALSE}
#| label: tbl-modelsummary
#| tbl-cap: "Summary of Key Model Estimates for Ontario, Federal, and Conservative/Liberal Models"
#| echo: false
#| message: false

library(modelsummary)

#### Read data ####
analysis_data <- read_csv("../data/02-analysis_data/analysis_data.csv")

#### Model for Ontario ####

# Filter data for Ontario
ontario_data <- analysis_data %>% 
  filter(region == "Ontario")

# Aggregate total donations per party per year in Ontario
total_donations_ontario <- ontario_data %>%
  group_by(political_party, donation_year) %>%
  summarize(
    Total_Donations = sum(amount, na.rm = TRUE),
    In_Power = first(recipient_in_power),
    Party_Size = first(Party_Size),
    Election_Year = first(Election_Year)
  ) %>%
  ungroup()

# Log-transform the dependent variable to stabilize variance
total_donations_ontario <- total_donations_ontario %>%
  mutate(Log_Total_Donations = log(Total_Donations + 1))  # Adding 1 to avoid log(0)

# Run Regression for Ontario with Controls
model_ontario <- lm(Log_Total_Donations ~ In_Power + Party_Size + Election_Year, data = total_donations_ontario)

#### Model for Federal ####

# Filter data for Federal
federal_data <- analysis_data %>% 
  filter(region == "Federal")

# Aggregate total donations per party per year at Federal level
total_donations_federal <- federal_data %>%
  group_by(political_party, donation_year) %>%
  summarize(
    Total_Donations = sum(amount, na.rm = TRUE),
    In_Power = first(recipient_in_power),
    Party_Size = first(Party_Size),
    Election_Year = first(Election_Year)
  ) %>%
  ungroup()

# Log-transform the dependent variable to stabilize variance
total_donations_federal <- total_donations_federal %>%
  mutate(Log_Total_Donations = log(Total_Donations + 1))  # Adding 1 to avoid log(0)

# Run Regression for Federal with Controls
model_federal <- lm(Log_Total_Donations ~ In_Power + Party_Size + Election_Year, data = total_donations_federal)

# Filter for Conservative and Liberal parties
cl_data <- analysis_data %>%
  filter(political_party %in% c(
    "Conservative Party of Canada",
    "Liberal Party of Canada",
    "Progressive Conservative Party of Ontario",
    "Liberal Party of Ontario"
  ))

# Create variables
cl_data <- cl_data %>%
  mutate(
    In_Power = recipient_in_power,  # Already binary (1/0)
    Party = ifelse(
      political_party %in% c("Conservative Party of Canada", "Progressive Conservative Party of Ontario"),
      1,  # Conservative
      0   # Liberal
    )
  )

# Aggregate total donations per party per year
total_donations_cl <- cl_data %>%
  group_by(political_party, donation_year) %>%
  summarize(
    Total_Donations = sum(amount, na.rm = TRUE),
    In_Power = first(In_Power),
    Party = first(Party),
    Election_Year = first(Election_Year)
  ) %>%
  ungroup()

# Log-transform the dependent variable to stabilize variance
total_donations_cl <- total_donations_cl %>%
  mutate(Log_Total_Donations = log(Total_Donations + 1))  # Adding 1 to avoid log(0)

# Run the regression with interaction term
model_cl <- lm(Log_Total_Donations ~ In_Power * Party + Election_Year, data = total_donations_cl)

# Define the list of models
models <- list(
  "Ontario" = model_ontario,
  "Federal" = model_federal,
  "Conservative/Liberal" = model_cl
)

# Define custom coefficient labels
coef_labels <- c(
  "(Intercept)" = "Intercept",
  "In_Power" = "In Power",
  "Party_Size" = "Party Size",
  "Election_Year" = "Election Year",
  "Party" = "Party (1=Conservative, 0=Liberal)",
  "In_Power:Party" = "In Power * Party"
)

# Generate the regression table without significance stars, rounded to 3 decimals
modelsummary(
  models, 
  coef_map = coef_labels, 
  stars = FALSE,
  fmt = 3,
  gof_omit = "AIC|BIC|Log.Lik.|F|RMSE|R2",
  output = "markdown"
)

```

## Ontario Results

In the Ontario model, Party_Size is a significant predictor of total
donations, with a coefficient of 2.956 (standard error = 0.485, p \<
0.001). This indicates that major parties in Ontario receive
significantly higher total donations than minor parties, highlighting
the influence of party prominence on fundraising success.

The In_Power variable has a coefficient of 0.884 (standard error =
0.540, p = 0.106), which is not statistically significant at the
conventional 5% level. This suggests that being the ruling party in
Ontario does not have a significant direct effect on the total amount of
donations received.

The Election_Year variable is also not significant in the Ontario model
(coefficient = 0.239, standard error = 0.436, p = 0.586), indicating
that donations are not substantially influenced by whether they occur
during an election year.

## Federal Results

In the Federal model, Party_Size remains a significant predictor, with a
coefficient of 3.450 (standard error = 0.259, p \< 0.001). This
reinforces the importance of party size in attracting donations at the
federal level.

The In_Power variable in the Federal model has a coefficient of 0.637
(standard error = 0.3882, p = 0.103), which is not statistically
significant. Similar to the Ontario context, being in power at the
federal level does not have a significant direct effect on the total
donations received.

The Election_Year variable is not significant in the Federal model
(coefficient = -0.0752, standard error = 0.2209, p = 0.734), suggesting
no substantial impact of election years on donation amounts at the
federal level.

## Conservative/Liberal Results

In the Conservative/Liberal model, In_Power has a significant positive
effect, with a coefficient of 1.202 (standard error = 0.497, p =
0.01827). Additionally, the interaction term In_Power × Party is
significantly negative, with a coefficient of -1.718 (standard error =
0.711, p = 0.01845). This interaction indicates that the effect of being
in power on total donations varies between the Conservative and Liberal
parties.

For the Conservative Party (Party = 1), the combined effect of being in
power is 1.2023 + (-1.7175) = -0.5152, suggesting that being in power is
associated with lower total donations for the Conservative Party. For
the Liberal Party (Party = 0), the effect of being in power remains
1.2023, indicating that being in power is associated with higher total
donations for the Liberal Party.

The Party variable itself is significant (coefficient = 1.3648, standard
error = 0.4968, p = 0.00774), suggesting that, overall, the Conservative
Party tends to receive more donations than the Liberal Party when not
considering the interaction effect.

The Election_Year variable is not significant in the
Conservative/Liberal model (coefficient = -0.0527, standard error =
0.3777, p = 0.88946).

```{r echo=FALSE, message=FALSE, warning=FALSE}
#| label: fig-estimates
#| fig-cap: "Coefficient Estimates for Ontario, Federal, and Conservative/Liberal Models"
#| echo: false
#| message: false

library(broom)
library(ggplot2)
library(dplyr)
# Tidy models and combine
tidy_all <- bind_rows(
  tidy(model_ontario) %>% mutate(Model = "Ontario"),
  tidy(model_federal) %>% mutate(Model = "Federal"),
  tidy(model_cl) %>% mutate(Model = "Conservative/Liberal")
)

# Select relevant terms and exclude intercepts
tidy_plot <- tidy_all %>%
  filter(term != "(Intercept)") %>%
  mutate(term = case_when(
    term == "In_Power" ~ "In Power",
    term == "Party_Size" ~ "Party Size",
    term == "Election_Year" ~ "Election Year",
    term == "Party" ~ "Party (1=Conservative, 0=Liberal)",
    term == "In_Power:Party" ~ "In Power * Party",
    TRUE ~ term  # Keeps other terms unchanged
  ))

# Plot coefficients with error bars
ggplot(tidy_plot, aes(x = term, y = estimate, color = Model)) +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), 
                width = 0.2, position = position_dodge(width = 0.5)) +
  labs(
    x = "Predictors",
    y = "Estimated Coefficients"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

@fig-estimates highlights the consistently strong positive effect of
Party_Size on total donations in both the Ontario and Federal models,
emphasizing that major parties attract significantly more donations than
minor parties. It also illustrates the significant interaction effect
between In_Power and Party in the Conservative/Liberal model,
demonstrating the differing impact of being in power on donations for
the two major parties.

In summary, the analyses reveal that party size is a significant
determinant of total donations at both provincial and federal levels,
while the effect of being in power on donations is contingent upon party
affiliation, particularly between the Conservative and Liberal parties.

# Discussion {#sec-discussion}

## Financial Influence Shapes Political Landscapes

At the heart of politics is finance, as revealed by Traag’s research on
the complex contagion of campaign donations in the United States
[@traag2016complex]. Traag’s study found that most campaign
contributions come from a small number of wealthy donors linked by large
social networks, demonstrating how a few individuals can shape political
outcomes. Similar dynamics appear to be at play in Canada. Here, donors
often prefer to support parties not currently in power. One reason might
be the hope that challengers will bring needed policy changes. More
realistically, as noted by Mollona [@mollona2022ethics], politicians,
especially those out of power, often seek funding in exchange for
promoting donors’ interests. Over time, this can lead to a pattern where
donors shift their support to whoever best serves their goals, fostering
continuous political competition. A key finding of this study is that at
both the Ontario and federal levels, the Conservative Party generally
attracts more financial backing than other parties. This steady flow of
resources may help the Conservatives maintain a strong presence in the
political arena, influencing agenda-setting and public debate. With more
funds, they can more easily finance voter outreach, shape their
messaging, and solidify their support networks. While this financial
advantage may help sustain dynamic election contests, it also
underscores concerns about whose interests gain the most traction. When
one party consistently out-raises others, the policy conversation may
lean toward those who provide the most money, potentially sidelining
issues important to groups with fewer financial resources.

## International Comparisons and the Importance of Regulation

These trends are not confined to Canada. In the United Kingdom, for
example, large private donations, often from very wealthy “superdonors”,
widen the gap between major parties and smaller competitors
[@draca2022financing]. Such patterns highlight a broader democratic
challenge, as money can concentrate power and influence political
direction in ways that may not always reflect broad public preferences.
Rather than serving purely as a resource that encourages healthy
competition, political donations can also become a tool that benefits
certain parties and voices more than others. Regulation and transparency
are essential. Clearer rules and better public reporting help citizens
understand who funds whom and why. Drawing lessons from other countries
may guide reforms aimed at limiting undue influence. Stronger oversight,
spending caps, and easy-to-access donor databases can empower voters to
make informed choices. Without these safeguards, the already powerful,
whether individuals, industries, or interest groups, can gain even more
sway, risking an erosion of trust in democratic institutions.

## Weaknesses and next steps

This study provides valuable insights into donation behaviours across
Canada but it its limitations. First, the analysis is applied to only
federal and provincial donation records, overlooking regional variations
found in other provinces and territories. The geographical limitations
of this study hinder its ability to generalize across the entire
country. Moreover, the study only spans from 2006 to 2024 which economic
or political events could have influence over the donation patterns that
the models do not account for. Additionally qualitative factors such as
personal beliefs, specific policy preferences and campaign effectiveness
are not directly addressed in this study. Data quality issues such as
potential missing values or inconsistencies could also have affected the
accuracy of the findings here. Lastly, the linear regression models here
assume a linear relationship between predictors and donations which can
oversimplify the complexities.

Future research should not only aim to address these limitations, but
expand the analysis to other provinces and territories across Canada.
This would enhance the generalizability of the results and provide a
more nuanced understanding of regional differences. Furthermore,
utilizing nonlinear modelling techniques or looking at a broader set of
predictor variables such as donor demographics or economic indicators
could better capture the nuances of donor behaviour and also mitigate
the risk of omitted variable bias. Comparative studies with other
democratic countries, beyond Canada, could provide more general
perspectives on how other political systems and regulatory frameworks
influence or are influenced by donation behaviours. Other valuable areas
of future exploration include assessing the impact of media coverage on
donation trends. Finally, by addressing these areas, research can build
on the current study's findings and contribute to a greater
understanding of political donations.

\newpage

\appendix

# Appendix {#sec-appendix}

## Appendix A: Data Cleaning Notes

Rigorous data cleaning was done to the dataset used in this
investigation to assure correctness and dependability. The raw donation
records were sourced from the Investigative Journalism Foundation (IJF)
which include political donations at both federal and provincial levels,
specifically Ontario, from 2013 to 2024. The details of the data
cleaning processed are described below.

First, missing and incomplete values were searched for in key variables
such as Amount Donated, Donation Year, Political Party and Region.
Excluding any records with missing values in these crucial fields to
minimise biases created from incomplete data entries and to maintain
data integrity.

Inconsistencies in reporting caused political party names to appear in
different formats. To fix this, party names were given constant and
standardized names across all the records. For example, if there was a
variation such as "Liberal Party of Ontario" and "Ontario Liberal Party"
these were combined into a single name "Ontario Liberal Party" to ensure
accurate attribution of donations to parties.

Fields that were also standardized were the donor information which
includes names and locations. Any typographical errors were corrected,
and formatting was made consistent, for example the structure "First
Name Last Name" instead of "Last Name, First Name". But, in cases where
standardization could introduce inaccuracies, original donor names were
kept to avoid dealing with individuals with similar names.

Two columns, (amount_monetary) and (amount_non_monetary) which
correspond to monetary and non-monetary contributes are included in the
data set. Using these a new variable was created that sums these two
fields together for every record. In doing this, the dataset now
provides a comprehensive measure of the total value of donations
received by a party. Moreover, monetary values were converted into a
consistent currency format to disregard any discrepancies in the
currency symbols or formatting and to ensure accurate aggregation and
proper comparison.

Individual donation records were aggregated in order to perform proper
analysis at the party-year level. For each political party within each
year donations were summed yielding a dataset where each entry
represented the total donations received by a party for a specific year.
Aggregating the data like this enabled the examination of trends over
time and impact assessment factors like power status and election years
on donation amounts.

Several binary variables were introduced to capture key aspects of the
data:

Recipient_in_Power: Indicates whether the recipient party was in power
during the donation year (1 for in power, 0 for not in power).
Party_Size: Categorizes parties as major (1) or minor (0) based on
legislative representation and historical influence. Election_Year:
Indicates whether the donation was made during an election year (1 for
election year, 0 for non-election year). These variables were imperative
for the regression analysis and allowed the assessment of the influence
of these on total donations.

The distribution of total donation amounts was highly skewed, with a
small number of large donations and many smaller ones. In order to
address this, a logarithmic transformation was applied to the
Total_Donations variable creating a new variable called
Log_Total_Donations. This transformation also helped with handling zero
values by adding a small constant before taking the logarithm.

Cross-validation was used against original records and official
financial statements to ensure accuracy of the cleaned dataset. This
step required spot-checking aggregated totals and individuals records to
confirm that the cleaning and aggregation process did not create more
errors.

There were challenges that occurred due to inconsistencies in the raw
data, these included typographical errors, reporting format errors and
missing donor types. Some limitations remained despite the effort to
standardize and correct these issues, due to the quality of the original
data source. When donor type information was missing and legal
restrictions indicated that only individuals could donate, the donor
type was inferred as "Individual". Despite this assumption being legally
reasonable in the context of the paper, it may not capture all nuances
in the data.

\newpage

## Appendix B: IJF Data Collection Methodology

The Investigative Journalism Foundation (IJF) employs a comprehensive
approach to collect and compile political donation data across Canada.
Understanding the IJF's data collection methodology is crucial for
assessing the reliability and scope of the data used in this analysis.

**Data Sources and Acquisition**

Every day, the IJF checks election agency websites at the federal level
and in each province and territory for new political donations data.
Historical data is obtained from a variety of government sources,
including provincial archives, legislative libraries, and elections
agencies, covering the period from 1993 to the present day, though this
time range varies by jurisdiction.

Political parties and candidates are legally required to submit records
of donations, which are maintained and made publicly accessible by
election agencies. The frequency and format of these disclosures differ
among jurisdictions; some require quarterly returns, while others
mandate annual reports. The IJF systematically gathers these records to
create a unified database of political donations.

**Scope of Data Collected**

The IJF's dataset encompasses donations to various political entities,
including:

-   Parties: Federal and provincial registered political parties.

-   Candidates: Individual candidates running for office.

-   Riding Associations: Also known as electoral district or
    constituency associations.

-   Leadership and Nomination Contestants: Individuals participating in
    party leadership races or seeking nomination as candidates.

The dataset includes records from all 13 provinces and territories, as
well as city-specific donations from municipalities like Victoria and
Vancouver.

**Donation Laws and Contribution Limits**

Jurisdictions have varying limits on the donation amount and types of
entities allowed to donate. For example, in the federal level, only
individuals who are Canadian citizens or permanent residents can donate,
with a maximum annual contribution amount. Comparatively, there are some
provinces that allow donations from corporations, organizations and
unions and may have different or no contribution limits.

**Data Formats and Conversion**

The source data is available in multiple formats, including downloadable
spreadsheets, PDFs, and HTML files. For jurisdictions where data is only
available in PDF format, the IJF employs Optical Character Recognition
(OCR) technology to convert the documents into machine-readable CSV
files. OCR tools such as Adobe Export PDF, Cometdocs, and ABBYY
FineReader are used for this purpose.

**Data Cleaning and Standardization**

Manual cleaning is extensively done to correct errors that arise from
OCR conversion and to standardize the data. The key steps are:

Correcting OCR Errors: Manual verification and correction of misread
characters (e.g., misinterpreted currency symbols or letters).
Standardizing Donor and Party Names: Ensuring consistency in naming
conventions for donors and political parties (e.g., standardizing
"Progressive Conservative Party of Ontario" across all records).
Formatting Dates and Amounts: Standardizing date formats to YYYY-MM-DD
and ensuring monetary values are consistently formatted. The IJF also
amalgamates similar categories in variables such as donor types to
improve data legibility (e.g., merging "Business" and "Corporation" into
a single "Corporation" category). In cases where donor type information
is missing and laws specify that only individuals can donate (e.g.,
Ontario after 2017), the IJF adds "Individual" to the donor type column.

**Limitations and Challenges**

Records submitted by political entities can contain typographical errors
or inaccuracies which are seen in the dataset. Inconsistent or
incomplete data can occur in reporting practices across different
jurisdictions. Moreover, OCR technology is imperfect at capturing
handwritings or poorly scanned documents introducing errors despite
manual correction.

## Appendix C: Idealized Methodology

The idealized methodology for analyzing how a political party's power
status affects the total donations it receives in Canada involves
conducting a comprehensive survey targeting individual political donors
across the country. This methodology aims to gather detailed data on
donor behaviours, motivations, and perceptions, providing deeper
insights into the factors influencing political donations at both
federal and Ontario provincial levels.

**Target Population**

The target population consists of individuals who have made financial
contributions to Canadian federal or provincial political parties,
candidates, riding associations, or leadership contestants between 2013
and 2024. This includes donors from all provinces and territories,
ensuring representation across different regions and political contexts.

**Sampling Method**

A stratified random sampling technique will be utilized to ensure
proportional representation from various demographics and regions. The
sample will be stratified according to:

-   Region: All provinces and territories.

-   Donor Type: Donors to parties in power versus opposition parties.

-   Demographics: Age, gender, income level, education level, and
    political affiliation.

In Ontario and other populous provinces, oversampling will be conducted
to capture more granular data and reduce the margin of error. This
approach ensures that key voter segments are adequately represented,
particularly those in regions with significant political activity.

**Sample Size and Response Rate**

To maintain a national margin of error of plus or minus 3% at a 95%
confidence level, a total sample size of approximately 1,500 respondents
is required. This sample size allows for meaningful analysis of
subgroups within the data. Given an estimated response rate of 10%,
outreach efforts will target approximately 15,000 potential respondents
across Canada.

**Recruitment Methods**

Respondents will be recruited using a combination of methods to maximize
reach and diversity:

-   Online Panel Providers: Collaborate with established online panel
    providers such as Ipsos, Leger, or Angus Reid to access a pool of
    verified respondents who have consented to participate in surveys.

-   Email Invitations: Where legally permissible and ethically
    acceptable, utilize contact information from public donor records to
    send email invitations directly to known political donors.

-   Social Media Advertising: Use targeted advertisements on platforms
    like Facebook, Twitter, and LinkedIn to reach potential respondents
    interested in political activities.

-   Telephone Outreach: Conduct telephone surveys using interactive
    voice response (IVR) systems and live calls to reach donors who may
    not be active online, particularly older individuals or those in
    rural areas.

To incentivize participation, all respondents will be entered into a
national sweepstakes with a chance to win one of ten \$500 cash prizes.

**Data Collection Methods**

Data collection will be conducted via:

-   Online Surveys utilizing platforms like Qualtrics or SurveyMonkey to
    administer the survey electronically. The survey will be designed to
    be mobile-friendly and accessible to individuals with disabilities.

-   Telephone Surveys for respondents preferring or requiring this
    method, trained interviewers will administer the survey, ensuring
    consistency with the online questionnaire.

The survey is designed to take approximately 10 to 15 minutes to
complete, minimizing drop-off rates while collecting comprehensive data.

**Data Validation Measures**

Several measures will be implemented to ensure data quality and
integrity:

-   Completion time checks will be used monitor the time respondents
    take to complete the survey to identify and exclude those who rush
    through it (speeding).

-   Implement reCAPTCHA technology to guard against bots and automated
    responses.

-   Include attention-check questions to identify inattentive or
    fraudulent responses.

-   After data collection, apply statistical weighting to adjust for any
    demographic discrepancies, ensuring the sample accurately reflects
    the population of political donors in Canada.

**Budget Allocation**

The budget for this study is allocated as follows:

-   Survey Development: \$8,000

    -   This covers costs associated with designing the survey,
        including question formulation, structuring for clarity and
        flow, pilot testing, and ensuring compliance with ethical
        standards. Additionally, funds account for bilingual translation
        services (English and French). This also covers the cost of the
        potential monetary gifts given after completing the survey.

-   Sampling and Recruitment: \$15,000

    -   Expenses related to accessing online panel respondents,
        purchasing targeted advertising on social media, and
        collaborating with panel providers. Funds also cover the
        services of statisticians to ensure the sampling aligns with the
        population distribution.

-   Respondent Incentives: \$20,000

    -   Allocated for compensating participants with \$10 electronic
        gift cards or funding the national sweepstakes prizes.
        Incentives are crucial for improving response rates and
        attracting a diverse respondent pool.

-   Data Collection: \$25,000

    -   Operational costs for administering the survey, including online
        platform fees, telephone survey expenses, interviewer training,
        and call center operations. This ensures data collection is
        efficient and reaches a broad audience.

-   Data Analysis and Reporting: \$12,000

    -   Funds dedicated to analyzing the collected data, including
        hiring data analysts proficient in statistical software,
        conducting comprehensive analyses, and preparing detailed
        reports and visualizations to communicate findings effectively.

-   Ethical Compliance and Data Security: \$5,000

    -   Covers the costs of obtaining ethical approvals, ensuring
        compliance with privacy legislation (e.g., Personal Information
        Protection and Electronic Documents Act), and implementing
        robust data security measures to protect respondent information.

-   Total Estimated Budget: \$85,000

**Timeline**

-   Months 1-2: Survey design, ethical approval processes, and pilot
    testing to refine the questionnaire.

-   Months 3-4: Sampling frame development and recruitment of
    respondents through various channels.

-   Months 5-6: Data collection via online and telephone surveys.

-   Month 7: Data cleaning, validation, and preliminary analysis.

-   Months 8-9: Comprehensive data analysis and report preparation.

-   Month 10: Dissemination of findings through academic publications
    and presentations.

#### Idealized Survey Design

This survey helps to gather data about those who donate to political
parties and acquire more qualitative details about a sample of the
population. The survey can be accessed using [this
link](https://docs.google.com/forms/d/e/1FAIpQLSdj2RxS9fjZ_n6Q54zjSyoJ7tAJT7ssuTF3V1tVk6a_Rhyp4w/viewform).

**Survey Structure**

**Introduction**

Thank you for participating in this important survey. Your responses
will help us understand how Canadians make decisions about donating to
political parties at both the federal and Ontario provincial levels. By
sharing your perspectives, you will contribute to a deeper understanding
of what motivates political donations, what factors shape donor
decisions, and how parties can improve their engagement with supporters.
Your answers are completely confidential and will be used solely for
research purposes.

**Section 1: Demographics**

State of Residence:

-   Dropdown list of all provinces and territories.

Age:

-   18-24

-   25-34

-   35-44

-   45-54

-   55-64

-   65+

Gender:

-   Male

-   Female

-   Non-binary/Other

-   Prefer not to say.

Annual Household Income Level:

-   Less than \$25,000

-   \$25,000-\$49,999

-   \$50,000 - \$74,999

-   \$75,000 - \$99,000

-   \$100,000 - \$149,000

-   \$150,000 or more

-   Prefer not to say

Highest Level of Education Completed:

-   Less than high school

-   High school diploma or equivalent

-   Some college, no degree

-   Associate degree

-   Bachelor's degree

-   Graduate or professional degree

Political Affiliation:

-   Conservative

-   Liberal

-   NDP

-   Green

-   Other (please specify)

-   No specific affiliation

**Section 2: Donation Behavior**

Have you ever donated to a political party?

-   Yes, currently donate

-   Yes, but not currently

-   No, never donated

How frequently do you donate to political parties?

-   Never

-   Rarely (once a year or less)

-   Occasionally (a few times a year)

-   Regularly (monthly or more)

Average Annual Donation Amount (to all political parties combined):

-   Less than \$200

-   \$200 - \$1,000

-   \$1,001 - \$5,000

-   \$5,001 - \$10,000

-   More than \$10,000

-   Prefer not to say

Preferred Donation Channels: (Select all that apply)

-   Online platforms (e.g., party websites, crowdfunding sites)

-   Mail-in donations

-   In-person events (e.g., fundraisers, rallies)

-   Phone solicitations

-   Other (please specify)

**Section 3: Motivations for Donating**

What motivates you to donate to a political party? (Select all that
apply)

-   Support for specific policies or issues

-   Desire to influence election outcomes

-   Loyalty to the party or its leaders

-   Social pressure or community influence

-   Tax benefits

-   Personal connection to a candidate

-   Other (please specify)

Which factors most influence your decision to donate? (Select all that
apply)

-   Candidate’s leadership qualities

-   Party’s stance on key issues

-   Party’s track record and past performance

-   Recommendations from peers, family, or community leaders

-   Media coverage and public opinion polls

-   Urgency or importance of the upcoming election

-   Other (please specify)

**Section 4: Impact of Campaigns and Events**

-   Have recent political events (e.g., elections, leadership changes,
    policy announcements) influenced your donation behavior?

    -   Yes

    -   No

    If yes, please specify the event(s) and how they influenced your
    donations:\
    *(Open-ended)*

-   How do you perceive the effectiveness of political campaigns in
    encouraging donations?

    -   Very ineffective

    -   Ineffective

    -   Neutral

    -   Effective

    -   Very effective

-   Have you ever stopped donating or reduced donations because of a
    party’s perceived performance or power status (e.g., losing seats,
    scandals)?

    -   Yes (please explain why) *(Open-ended)*

    -   No

**Section 5: Feedback and Suggestions**

-   In your opinion, what could political parties do to encourage more
    donations from supporters?\
    *(Open-ended)*

-   Any additional comments or suggestions regarding political
    donations?\
    *(Open-ended)*

**Thank you for participating!** Your feedback is valuable and will help
researchers and political parties understand the dynamics of political
giving in Canada. Your insights contribute to more informed policies and
practices, ultimately shaping a more responsive and transparent
political landscape.

\newpage

# References